Project Light

### 2015-09-08 paper/paper.tex
add figure with screenshots
add keywords
minor changes in whole paper and notes

In the last decades, the rise embedded systems changed daily lives. % Source?
However, these systems often had limited networking capabilities or at least used their own, proprietary and thus unaccessible network technology.
Because of the huge variety of valuable possibilities, the impact of embedded systems will be advanced even further within the IoT.

@@

Another approach is to pre-initialize the keys in the factory.
This would mean that either there is a small set of assigned keys, which would lead to vulnerabilities to brute-force attacks; or each mote gets an individual key, which requires a secure transmission and attribution of keys to motes\footnote{This would also increase cost of motes, but they need to be as cheap as possible in order to be be affordable in great numbers.}.
Anyway, pre-initialization requires somebody other than the user/installer to know the keys, which diminishes security.

This paper proposes an alternative way to transmit initial keying material via light.
It is mainly targeted on private users without technical background and should provide them with a simple though secure process for initializing.

@@

There are several advantages of this method.
Light sensors are smaller and less misusable than jacks.
The physics of light is well understood by non-technical people, in contrast to radio waves.
Thus, leaking of the wireless sent secrets is less likely.
Moreover, multiple devices could be programmed concurrently by just placing them next to each other.

We implemented prototype for an Android smart phone and a Contiki\footnote{http://www.contiki-os.org/} mote.
Contiki is an open source, state-of-the-art operating system for IoT devices.
For Android phones, we implemented two versions of a light emitting application.
We also tested some techniques to make the transmission error-prone.
Besides, we examined how fast the whole transmission could become while maintaining a high chance of correctness.

@@

The target of the initialization phase is to tell the mote, when the actual data starts.
The communication between the mote and the phone is one-way.
That means, the mote cannot communicate with the Android app.
Therefore, the app has to guess, when the mote has successfully done the calibration and synchronization.
This can either be done by using a fixed time (let's say the app stays in calibration/synchronization phase for 10 seconds), or by user interaction (the user presses a button, when the mote indicates that it's ready).

@@

After successful synchronization, the mote waits for the initialization pattern.
It does that by reading single bits and comparing the last 8 bits (for ASCII encoding) to the pre-defined pattern.
Once the init pattern has been detected, the mote imediately switches to the data sending phase.

@@

In the data sending phase, we can finally transmit the data packet.
As we can see in Figure~\ref{fig:data_packet}, the packet contains the data length, the actual data and a checksum used for validation of the transmission in the validation phase~\ref{sub:validation}.

@@

Project Light consists of an Android app that creates and sends the key material and a Contiki process that runs on the TelosB sky mote and receives and validates it.
Both follow the aforementioned communication protocol very closely, but increase reliability and fail-safeness through the following design decisions.

@@

As the android timer that triggers the bit emission cannot be too early but in fact tends to be a little late\cite{mongia2010reliable}, the time at which the mote reads a bit is not in the middle of a period as suggested by Figure~\ref{fig:read_a_bit}.
It is at the very end of a period to maximize error avoidance due to timer delays.

@@

We have developed two different versions of an Android application for users to enter keys and transmit them.

The first version is written using web technologies (i.e. HTML, CSS, and JavaScript) and deployed to the smart phone using Apache~Cordova\footnote{https://cordova.apache.org/}.
Because of this, the application can also run in most desktop browsers and can easily be ported to work with other smart phone operating systems, too, e.g. iOS or Windows Phone 8.1.
In this version, almost all parameters for the transmission are adjustable by the user: length of one bit, length of initialization phase, initialization-pattern, start-pattern, if hamming-encoding should be used.
Therefore it is convenient to experiment with various settings and transmission parameters and to test of the Contiki process.
There are two disadvantages of this implementation, which motivated us to build a second edition:
First, experiments using the flash light for light emission showed, that there are no predictable timing behavior, i.g. the flash light was simply to slow to allow short transmission times.
This forced us to use the screen for light emission, which resulted in far smaller brightness margins to be read by the mote.
Secondly, in order to get reliable transmission quality, we need to become as close to real-time as possible.
This technology stack indicates bad real-time behavior, i.g. the Cordova runtime and the JavaScript interpreter could be omitted.

%figure: both screenshots next to each other

Hence, we built a native Android application written in Java.
Still on Android, it is impossible to guarantee real-time behavior, but we may get smaller delays without JavaScript involved.
The Java implementation also benefits from the ability of the native Java timer to readjust itself, i.g. if one event is fired too late such as wrong bit gets transmitted, the delay does not add up for the next bit.
This comes hand-in-hand with the Hamming(8,4)-encoding we used, because this could fix potential one-bit-errors in each byte.
Another advantage of this version is is the flash light.
Depending on the smart phone model, the flash light was as fast as screen flickering even for small bit lengths, while being much brighter.
Especially when attempting to program multiple motes at once this comes very handy.
To improve usability and discoverability, we dropped controls for some parameters, which led to a much leaner design as shown in figure~\ref{fig:screenshots}.

@@

On the mote the communication protocol is implemented in a Contiki process.
The Contiki operating system offers a real-time timer called \textit{rtimer} which eliminates timer inaccuracies on the mote side.
The mote uses its light sensor to sense the current brightness.
Once the process is started an LED-countdown (red -> blue -> green) indicates the start of the key initialization procedure.
A \mbox{\textit{2-means}} algorithm is used to create two clusters of \textit{bright} and \textit{dark} values and to compute their mean values.
Subsequently a new value is classified as \textit{0} or \textit{1} depending on which mean value is closer to it.
After the successful synchronization and initialization a blue LED indicates that the mote is receiving data.
If an error is detected by the Hamming-code or if the computed checksum does not match the transmitted one, a red LED indidcates the failure and the process should be restarted.
A green LED shows that the key material has successfully been received and was saved to the permanent flash? memory of the mote.

@@

As seen in (code example), calls to the driver that are being issued before the key material initialization are actually being canceled right now, but should eventually just be delayed and reissued once the initialization was finished.
Including the light app on a mote automatically redefines \textit{NETSTACK\_CONF\_LLSEC}\footnote{This global constant defines the link layer driver that is used by the mote.} to point to our link layer driver wrapper (and no further configuration needs to be done?).

@@

Future work should target the further automation of the presented process, by establishing a bidirectional communication.
Using its camera the mobile phone could detect the current LED-state of the mote and react accordingly, e.g. start the data transmission or restart the whole process.
This would also make it easier to initalize several motes at once, which because of lack of synchronization is currently quite tedious.

User-friendliness should be prioritized as success and adoption of this presented additional security measure will greatly benefit from it.

>>>>

In the last decades, the rise of embedded systems changed daily lives. % Source?
However, these systems often had limited networking capabilities or at least used their own, proprietary and thus unaccessible network technology.
Because of the huge variety of valuable possibilities, the impact of embedded systems will be advanced even further within the IoT.

@@

Another approach is to pre-initialize the keys in the factory.
This would mean that either there is a small set of assigned keys, which would lead to vulnerabilities to brute-force attacks; or each mote gets an individual key, which requires a secure transmission and attribution of keys to motes\footnote{This would also increase cost of motes, but they need to be as cheap as possible in order to be affordable in great numbers.}.
Anyway, pre-initialization requires somebody other than the user/installer to know the keys, which diminishes security.

This paper proposes an alternative way, which is transmitting initial keying material via light.
It is mainly targeted on private users without technical background and should provide them with a simple though secure process for initializing.

@@

There are several advantages of this method.
Light sensors are smaller and less misusable than jacks.
The physics of light is well understood by non-technical people, in contrast to radio waves.
Thus, leaking of the wireless sent secrets is less likely.
Moreover, multiple devices could be initialized concurrently by just placing them next to each other.

We implemented a prototype for an Android smart phone and a mote running the Contiki OS\footnote{http://www.contiki-os.org/}.
Contiki is an open source, state-of-the-art operating system for IoT devices.
We also tested some techniques to make the transmission error-prone.
Besides, we examined how fast the whole transmission could become while maintaining a high chance of correctness.

@@

The target of the initialization phase is to tell the mote, when the actual data starts.
The communication between the mote and the phone is one-way.
That means, the mote cannot communicate with the Android app.
Therefore, the app has to guess, when the mote has successfully done calibration and synchronization.
This can either be done by using a fixed time (let's say the app stays in calibration and synchronization phase for 10 seconds), or by user interaction (the user presses a button, when the mote indicates that it's ready).

@@

After successful synchronization, the mote waits for the initialization pattern.
It does that by reading single bits and comparing the last 8 bits (for ASCII encoding) to the pre-defined pattern.
Once the init pattern has been detected, the mote immediately switches to the data sending phase.

@@

In the data sending phase, we can finally transmit the data packet.
As we can see in Figure~\ref{fig:data_packet}, the packet contains the data length, the actual data and a checksum used for validation of the received data in the validation phase~\ref{sub:validation}.

@@

Project Light consists of an Android app that creates and sends the keying material and a Contiki process that runs on the TelosB sky mote and receives and validates it.
Both follow the aforementioned communication protocol very closely, but increase reliability and fail-safeness through the following design decisions.

@@

As the android timer that triggers the bit emission cannot be too early but in fact tends to be a little late\cite{mongia2010reliable}, the time at which the mote reads a bit is not in the middle of a period as suggested by Figure~\ref{fig:read_a_bit}.
It is at the very end of a period to maximize error avoidance due to timer delays\footnote{The Contiki timer that triggers the reading of the bit is a real-time timer.}.

@@

We have developed two different versions of the Android application for users to enter keys and transmit them.

% Ich wuerde den gesamten folgenden Abschnitt weglassen und gar nicht so richtig ueber die js timer Sachen reden. Vielleicht lieber noch einen kurzen Abschnitt generell ueber timing Sachen?

The first version is written using web technologies (i.e. HTML, CSS, and JavaScript) and deployed to the smart phone using Apache~Cordova\footnote{https://cordova.apache.org/}.
Because of this, the application can also run in most desktop browsers and can easily be ported to work with other smart phone operating systems, too, e.g. iOS or Windows Phone 8.1.
In this version, almost all parameters for the transmission are adjustable by the user: length of one bit, length of initialization phase, initialization-pattern, start-pattern, if hamming-encoding should be used.
Therefore it is convenient to experiment with various settings and transmission parameters and to test of the Contiki process.
There are two disadvantages of this implementation, which motivated us to build a second edition:
First, experiments using the flash light for light emission showed, that there are no predictable timing behavior, e.g. the flash light was simply to slow to allow short transmission times.
This forced us to use the screen for light emission, which resulted in far smaller brightness margins to be read by the mote.
Secondly, in order to get reliable transmission quality, we need to become as close to real-time as possible.
This technology stack indicates bad real-time behavior, i.g. the Cordova runtime and the JavaScript interpreter could be omitted.

% bis hier hin weg

% \begin{figure}
% 	\centering
% 	\includegraphics[scale=.35]{images/screenshots.png}
% 	\caption{The two Android App versions side-by-side (web on the left, native on the right).}
% 	\label{fig:screenshots}
% \end{figure}

% Und den dann natuerlich noch anpassen

Hence, we built a native Android application written in Java.
Still on Android, it is impossible to guarantee real-time behavior, but we may get smaller delays without JavaScript involved.
The Java implementation also benefits from the ability of the native Java timer to re-adjust itself, which means that timing errors do not add up.
In that way, the phone and the mote do not get out of sync, no matter how long the transmission takes.
% , e.g. if one event is fired too late such as wrong bit gets transmitted, the delay does not add up for the next bit.
This comes hand-in-hand with the Hamming(8,4)-encoding we used, because this could fix potential one-bit-errors in each byte.
Another advantage of this version is is the flash light.
Depending on the smart phone model, the flash light was as fast as screen flickering even for small bit lengths, while being much brighter.
Especially when attempting to program multiple motes at once this comes very handy.
To improve usability and discoverability, we dropped controls for some parameters, which led to a much leaner design as shown in figure~\ref{fig:screenshots}.

@@

On the mote, the communication protocol is implemented in a Contiki process.
The Contiki operating system offers a real-time timer called \textit{rtimer} which eliminates timer inaccuracies on the mote side.
The mote uses its light sensor to sense the current brightness.
Once the process is started, an LED-countdown (red -> blue -> green) indicates the start of the key initialization procedure.
A \mbox{\textit{2-means}} algorithm is used to create two clusters of \textit{bright} and \textit{dark} values and to compute their mean values~(\ref{sub:calibration}).
Subsequently a new value is classified as \textit{0} or \textit{1} depending on which mean value is closer to it.
After the successful calibration, synchronization and initialization a blue LED indicates that the mote is receiving data.
If an error is detected by the Hamming-code or if the computed checksum does not match the transmitted one, a red LED indidcates the failure and the process should be restarted.
A green LED shows that the key material has successfully been received and was saved to the permanent flash memory of the mote.

@@

As seen in (code example), calls to the driver that are being issued before the key material initialization are actually being canceled right now, but should eventually just be delayed and reissued once the initialization has finished.
Including the light app on a mote automatically redefines \textit{NETSTACK\_CONF\_LLSEC}\footnote{This global constant defines the link layer driver that is used by the mote.} to point to our link layer driver wrapper (and no further configuration needs to be done?).

@@

Future work should target the further automation of the presented process, by establishing a bidirectional communication.
Using its camera the mobile phone could detect the current LED-state of the mote and react accordingly, e.g. start the data transmission or restart the whole process.

To simplify the initialization of several motes at once, two things need be done.
First, the app should stay in calibration and synchronization phase, until the user presses a button.
Second, the mote should indicate via an LED that it reached the initialization phase~(\ref{sub:initialization}).
With those two concepts, the key transmission to several motes could work as follows: 
The user lays out all motes in a grid next to each other.
Then the phone is placed on top of them and the app is started.
All motes are placed, so that they receive the flickering of the phones flash light.
The motes are started one after another.
Once the motes are running, the user waits for all of them to indicate via an LED, that they reached the initialization phase.
They are now all waiting for the initialization pattern.
The user presses a button in the app and the phone sends the initialization pattern followed by the data packet.
When the transmission is over, the user can see via a red LED which motes failed to receive the data correctly.
The process needs to be restarted with those motes.

% den Satz verstehe ich nicht 
User-friendliness should be prioritized as success and adoption of this presented additional security measure will greatly benefit from it.
